---
title: "Hybrid-EP + Junction Tree Demo"
author: "Eric Chuu"
date: "4/27/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Computing the log normalizing constant of the G-Wishart density

Consider the G-Wishart density, which has the following form:

$$ f\left(\Omega \mid G \right) \propto \, | \Omega |^{(\delta-2)/2} \exp \left( - \mbox{tr} \left(\Omega \Lambda \right) /2 \right), $$
Here, $\Omega, \Lambda$ are $p \times p$ non-negative definite matrices, and $\delta > 2$ is the degrees of freedom. The normalizing constant is typically intractable, and the algorithm that we develop in this package targets this quantity.


### Case 1: $p = 30$. 

Define the following problem settings. We first randomly initialize the $30 \times 30$ adjacency matrix representation of the graph $G$. 

```{r, message = FALSE, warning = FALSE}
library(graphml)
set.seed(1234)
p = 30
Adj = matrix(rbinom(p^2,1,0.15), p, p)
Adj = Adj + t(Adj)
diag(Adj) = 0
Adj[Adj==1]=0
Adj[Adj==2]=1
diag(Adj) = 1
```

Next, we define the following degrees of freedom and scale matrix.

```{r}
b = 500
Y = matrix(rnorm(p*500), nrow = 500, ncol = p)
Lambda  = t(Y)%*%Y
```

Next, we compute the edge matrix which is needed for the junction tree part of the Hybrid-EP + JT algorithm.

```{r}
# compute the edge matrix
EdgeMat = graphml::getEdgeMat(Adj)
```


Finally, we compute the Hybrid-EP + JT appoximation to the log normalizing constant.

```{r}
start_time <- Sys.time()
hyb_approx = graphml::hybridJT(Adj, EdgeMat, b, Lambda, 1000)
end_time <- Sys.time()
hyb_approx
end_time - start_time
```

We compare this to Atay's method. 

```{r}
start_time <- Sys.time()
atay_approx = BDgraph::gnorm(Adj, b, Lambda, 1000)
end_time <- Sys.time()
atay_approx
end_time - start_time
```

We observe that the two estimates are quite close to each other, and that the Hybrid-EP + JT method is faster. 

### Case 2: $p = 60$. 

```{r, message = FALSE, warning = FALSE}
library(graphml)
set.seed(1234)
p = 30
Adj = matrix(rbinom(p^2,1,0.15), p, p)
Adj = Adj + t(Adj)
diag(Adj) = 0
Adj[Adj==1]=0
Adj[Adj==2]=1
diag(Adj) = 1
b = 500
Y = matrix(rnorm(p*500), nrow = 500, ncol = p)
Lambda  = t(Y)%*%Y
# compute the edge matrix
EdgeMat = graphml::getEdgeMat(Adj)
```


Finally, we compute the Hybrid-EP + JT approximation to the log normalizing constant.


```{r}
start_time <- Sys.time()
hyb_approx = graphml::hybridJT(Adj, EdgeMat, b, Lambda, 1000)
end_time <- Sys.time()
hyb_approx
end_time - start_time
```

We compare this to Atay's method. 

```{r}
start_time <- Sys.time()
atay_approx = BDgraph::gnorm(Adj, b, Lambda, 1000)
end_time <- Sys.time()
atay_approx
end_time - start_time
```


We observe that Atay's method fails to give a finite estimate, while the Hybrid-EP + JT quickly produces a sensible approximation.


